---
title: "Descriptive stats DHS data"
author: "Natalia Mejia"
date: "2/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
#Required packages

library(dplyr)
library(haven)
library(xtable)
library(plm)
library(pglm)
library(broom)
library(knitr) 
library(tidyverse)
library(tidyr)
library(haven)
library(stargazer)
library(kableExtra)
library(knitr)
library(expss)
library(forcats)
library(devtools)
library(questionr)
library(crunch)
library(table1)

```

```{r}

setwd("C:/Users/natim/Desktop/Thesis/Thesis2/Data/DHS")

#Getting final data

save(final_data, file = "final_data.RData")

load("final_data.RData")

```

```{r}
#Recoding variables

#Mom marital status

final_data$mom_marital <- factor(final_data$mom_marital)

final_data$mom_marital <- fct_collapse(final_data$mom_marital,
                                    "1-Married" = c("1-Married"),
                                    "2-Living with partner" = c("2-Living with partner"),
                                    "3-Single" = c("3-Separated", "4-Divorced", "6-Never in union"),
                                    "4-Other" = c("5-Widowed", "7-Dead", "8-Does not know"))
                                   
levels(final_data$mom_marital)

#Dad marital status

final_data$dad_marital <- factor(final_data$dad_marital)

final_data$dad_marital <- fct_collapse(final_data$dad_marital,
                                    "1-Married" = c("1-Married"),
                                    "2-Living with partner" = c("2-Living with partner"),
                                    "3-Single" = c("3-Separated", "4-Divorced", "6-Never in union"),
                                    "4-Other" = c("5-Widowed", "7-Dead", "8-Don´t know"))


levels(final_data$dad_marital)

#Education head of household

final_data$educ_head <- as.factor(final_data$educ_head)


final_data$educ_head <- fct_collapse(final_data$educ_head,
                                    "0-No education" = c("0-No education", "8-Don´t know"),
                                    "1-Primary" = c("1-Primary"),
                                    "2-Secondary" = c("2-Secondary"),
                                    "3-Higher" = c("3-Higher"))
final_data <- final_data %>%
  filter(!is.na(educ_head))


summary(final_data$educ_head)

```


```{r}
#Reason living school for context

label(final_data$reason_school) <- "Reason leaving school"

reason_school <- subset(final_data, final_data$reason_school != "No info")


table1(~ reason_school, 
       data= reason_school,
       caption = "Table XX. Reasons for dropping out of school",
       footnote = "Source: DHS Colombia 2015, own estimates")

```


```{r}
#Summary statistics 

label(final_data$dropouts) <- "School dropouts"
label(final_data$gender) <- "Gender"
label(final_data$age) <- "Age"
label(final_data$region) <- "Region"
label(final_data$resi) <- "Type of residence"
label(final_data$wealth) <- "Wealth status"
label(final_data$mom_marital) <- "Mothers´ marital status"
label(final_data$dad_marital) <- "Dad marital status"
label(final_data$educ_head) <- "Level of education head household"


table1(~dropouts + gender + age + region + resi + wealth + mom_marital + educ_head, 
       data= final_data,
       caption = "Table 2. Summary statistics main variables",
       footnote = "Source: DHS Colombia 2015, own estimates")
```

```{r}
#Cross tabulation education head and school status

head_school <- prop.table(table(final_data$educ_head, final_data$school_status),1)*100

head_school %>%
  kable(caption = "Cross tabulation head of household education level and school status of the children, column %",
        col.names = c("Dropout = 0", "Attending school = 1"),
        align = "c") %>%
kable_classic(full_width = F, html_font = "Cambria") 

```

```{r}
#Cross tabulation type of residence and school status 

resi_school <- prop.table(table(final_data$resi, final_data$school_status),1)*100

resi_school %>%
  kable(caption = "Cross tabulation type of residence and school status of the children, column %",
        col.names = c("Dropout = 0", "Attending school = 1"),
        align = "c") %>%
kable_classic(full_width = F, html_font = "Cambria") 
```
```{r}
#Cross tabulation wealth and school status 

wealth_school <- prop.table(table(final_data$wealth, final_data$school_status),1)*100

wealth_school %>%
  kable(caption = "Cross tabulation wealth status and school status of the children, column %",
        col.names = c("Dropout = 0", "Attending school = 1"),
        align = "c") %>%
kable_classic(full_width = F, html_font = "Cambria") 
```

```{r}
#Cross tabulation mom marital status and school status 

mom_school <- prop.table(table(final_data$mom_marital, final_data$school_status),1)*100

mom_school %>%
  kable(caption = "Cross tabulation mom marital status and school status of the children, column %",
        col.names = c("Dropout = 0", "Attending school = 1"),
        align = "c") %>%
kable_classic(full_width = F, html_font = "Cambria") 
```
```{r}
#Cross tabulation dad marital status and school status 

dad_school <- prop.table(table(final_data$dad_marital, final_data$school_status),1)*100

dad_school %>%
  kable(caption = "Cross tabulation dad marital status and school status of the children, column %",
        col.names = c("Dropout = 0", "Attending school = 1"),
        align = "c") %>%
kable_classic(full_width = F, html_font = "Cambria") 
```

```{r}
#Cross tabulation region and dropouts

region <- prop.table(table(final_data$region, final_data$dropouts),1)*100

region %>%
  kable(caption = "Cross tabulation region and dropouts, column %",
  
        align = "c") %>%
kable_classic(full_width = F, html_font = "Cambria") 
```






```{r}
#Creating table cross tabulations dep and indep variables


head_edu <- data.frame("Variables" =
                            c("0-No education", 
                              "1-Primary", 
                              "2-Secondary",
                              "3-Higher"),
                          "Dropout" = c(11.1, 9.5,8.2,6.9),
                          "Attending school" = c(88.9,90.5,91.7,93.1))

wealth <- data.frame("Variables" = 
                            c("Poorest", 
                              "Poor", 
                              "Middle",
                              "Rich",
                              "Richest"),
                          "Dropout" = c(11.3,9.7,7.8,5.6,3.2),
                          "Attending school"=c(88.7,90.2,92.1,94.4,96.7))


mom <- data.frame("Variables" =
                             c("1-Married",
                               "2-Living with partner",
                               "3-Single",
                               "4-Other"),
                           "Dropout" = c(7,10.1,8,10.4),
                           "Attending school" = c(93,89.9,92,89.5))

dad <- data.frame("Variables" =
                             c("1-Married",
                               "2-Living with partner",
                               "3-Single",
                               "4-Other"),
                           "Dropout" = c(20.6, 47.2, 16.5, 15.6),
                           "Attending school" = c(28.1, 43.3, 14.5, 14.1))


DATA <- rbind(head_edu, wealth, mom)

kbl(DATA, 
    booktabs = T,
    caption = "Table 3. Cross tabulations dependent and independent variables, column %",
    footnote = "Source: DHS Colombia 2015, own estimates") %>%
  pack_rows("Level of education head household", 1, 4) %>%
  pack_rows("Wealth status", 5, 9) %>%
  pack_rows("Mothers´ marital status", 10, 13) %>%
  kable_classic(full_width = F, html_font = "Cambria") 

```

```{r}
#Model 1 education head of household and covariate 1

model1 <- glm(dropouts ~ educ_head, data = final_data, family = "binomial")

summary(model1)

exp(coef(model1))
```

```{r}
#Model 1 education head of household and covariate 1

model1_cov1 <- glm(dropouts ~ educ_head + gender, data = final_data, family = "binomial")

summary(model1_cov1)

exp(coef(model1_cov1))
```

```{r}

#Model 1 education head of household and covariate 2

model1_cov2 <- glm(dropouts ~ educ_head + gender + resi, data = final_data, family = "binomial")

summary(model1_cov2)

exp(coef(model1_cov2))


```

```{r}
#Model 1 education head of household and covariate 3

model1_cov3 <- glm(dropouts ~ educ_head + region, data = final_data, family = "binomial")

summary(model1_cov3)

exp(coef(model1_cov3))
```
```{r}
#Model 1 education head of household and all covariates

model1_all <- glm(dropouts ~ educ_head + resi + gender + region, data = final_data, family = "binomial")

summary(model1_all)

exp(coef(model1_all))

```

```{r}
#Getting odd ratios and confidence intervals of each model to plot them with stargazer

#Model 1
OR.vector <- exp(coef(model1))
CI.vector <- exp(confint(model1))

p.values <- list(summary(model1)$coefficients[,4])


#Model 1 cov 1
OR.vector1 <- exp(coef(model1_cov1))
CI.vector1 <- exp(confint(model1_cov1))

p.values1 <- list(summary(model1_cov1)$coefficients[,4])

#Model 1 cov 2
OR.vector2 <- exp(coef(model1_cov2))
CI.vector2 <- exp(confint(model1_cov2))

p.values2 <- list(summary(model1_cov2)$coefficients[,4])

#Model 1 cov 3
OR.vector3 <- exp(coef(model1_cov3))
CI.vector3 <- exp(confint(model1_cov3))

p.values3 <- list(summary(model1_cov3)$coefficients[,4])

#Model 1 cov all
OR.vector4 <- exp(coef(model1_all))
CI.vector4 <- exp(confint(model1_all))

p.values4 <- list(summary(model1_all)$coefficients[,4])

```
```{r}
#Plotting

stargazer(model1, model1_cov1, model1_cov2,model1_cov3, model1_all, coef = list(OR.vector, OR.vector1, OR.vector2, OR.vector3, OR.vector4), 
          title = "Table 4. Stepwise model estimates school dropouts and education head of the household, estimates in odd ratios",
          dep.var.labels = "School dropouts",
          covariate.labels = c("Educ head-Primary",
                               "Educ head-Secondary",
                               "Educ head -Higher",
                               "Gender-Female",
                               "Type residence-Rural",
                               "Region-Oriental",
                               "Region-Bogota",
                               "Region-Central",
                               "Region-Pacific",
                               "Region-Amazonia"),
          ci = T,
ci.custom = list(CI.vector, CI.vector1, CI.vector2, CI.vector3, CI.vector4), single.row = F, type = "text", out = "model1.htm",
notes= c("Reference category Educ head-No education. Reference category Gender: Male. Reference type of residence: Urban.Reference category Region: Atlantic. Source: DHS Colombia 2015, own estimates"),
          notes.align = "l",
          notes.append = FALSE,
p=
c(p.values, p.values1, p.values2, p.values3, p.values4))


```


```{r}
#Model 2 socioeconomic status of the household 

model2 <- glm(dropouts ~ wealth , data = final_data, family = "binomial")

summary(model2)

exp(coef(model2))

```

```{r}
#Model 2 socioeconomic status of the household and gender

model2_cov1 <- glm(dropouts ~ wealth + gender, data = final_data, family = "binomial")

summary(model2_cov1)

exp(coef(model2_cov1))

```


```{r}
#Model 2 socioeconomic status of the household and type of residence

model2_cov2 <- glm(dropouts ~ wealth + resi, data = final_data, family = "binomial")

summary(model2_cov2)

exp(coef(model2_cov2))
```


```{r}
#Model 2 socioeconomic status of the household and region

model2_cov3 <- glm(dropouts ~ wealth + region, data = final_data, family = "binomial")

summary(model2_cov3)

exp(coef(model2_cov3))

```


```{r}
#Model 2 socioeconomic status of the household and all covariates

model2_all <- glm(dropouts ~ wealth + gender + resi + region, data = final_data, family = "binomial")

summary(model2_all)

exp(coef(model2_all))
```

```{r}
#Getting odd ratios and confidence intervals of each model to plot them with stargazer

#Model 2
OR.vector_m2 <- exp(coef(model2))
CI.vector_m2 <- exp(confint(model2))

p.values_m2 <- list(summary(model2)$coefficients[,4])

#Model 2 cov 1
OR.vector1.1 <- exp(coef(model2_cov1))
CI.vector1.1 <- exp(confint(model2_cov1))

p.values1.1 <- list(summary(model2_cov1)$coefficients[,4])

#Model 1 cov 2
OR.vector2.2 <- exp(coef(model2_cov2))
CI.vector2.2 <- exp(confint(model2_cov2))

p.values2.2 <- list(summary(model2_cov2)$coefficients[,4])

#Model 1 cov 3
OR.vector3.3 <- exp(coef(model2_cov3))
CI.vector3.3 <- exp(confint(model2_cov3))

p.values3.3 <- list(summary(model2_cov3)$coefficients[,4])

#Model 1 cov all
OR.vector4.4 <- exp(coef(model2_all))
CI.vector4.4 <- exp(confint(model2_all))

p.values4.4 <- list(summary(model2_all)$coefficients[,4])

```

```{r}
#Plotting

stargazer(model2, model2_cov1, model2_cov2,model2_cov3, model2_all, coef = list(OR.vector_m2, OR.vector1.1, OR.vector2.2, OR.vector3.3, OR.vector4.4), 
          title = "Table 5. Stepwise model estimates school dropouts and wealth status, estimates in odd ratios",
          dep.var.labels = "School dropouts",
          ci = T,
ci.custom = list(CI.vector_m2, CI.vector1.1, CI.vector2.2, CI.vector3.3, CI.vector4.4), single.row = F, type = "text", out = "model2.htm",
notes= c("Reference category Wealth stauts: Poorest. Reference category Gender: Male. Reference type of residence: Urban.Reference category Region: Atlantic. Source: DHS Colombia 2015, own estimates"),
 covariate.labels = c("Wealth status-Poor",
                               "Wealth status-Middle",
                               "Wealth status-Rich",
                      "Wealth status-Richest",
                               "Gender-Female",
                               "Type residence-Rural",
                               "Region-Oriental",
                               "Region-Bogota",
                               "Region-Central",
                               "Region-Pacific",
                               "Region-Amazonia"),
          notes.align = "l",
          notes.append = FALSE,
p=
c(p.values_m2, p.values1.1, p.values2.2, p.values3.3, p.values4.4))


```
```{r}
#Model 3 mothers´ marital status

model3 <- glm(dropouts ~ mom_marital , data = final_data, family = "binomial")

summary(model3)

exp(coef(model3))

```

```{r}
#Model 3 mothers´ marital status and gender 

model3_cov1 <- glm(dropouts ~ mom_marital + gender, data = final_data, family = "binomial")

summary(model3_cov1)

exp(coef(model3_cov1))

```

```{r}
#Model 3 mothers´ marital status and type of residence 

model3_cov2 <- glm(dropouts ~ mom_marital + resi, data = final_data, family = "binomial")

summary(model3_cov2)

exp(coef(model3_cov2))

```

```{r}
#Model 3 mothers´ marital status and type of residence 

model3_cov3 <- glm(dropouts ~ mom_marital + region, data = final_data, family = "binomial")

summary(model3_cov3)

exp(coef(model3_cov3))
```

```{r}
#Model 3 mothers´ marital status and type of residence 

model3_all <- glm(dropouts ~ mom_marital +gender + resi + region, data = final_data, family = "binomial")

summary(model3_all)

exp(coef(model3_cov3))
```


```{r}
#Getting odd ratios and confidence intervals of each model to plot them with stargazer

#Model 3
OR.vector_m3 <- exp(coef(model3))
CI.vector_m3 <- exp(confint(model3))

p.values_m3 <- list(summary(model3)$coefficients[,4])

#Model 2 cov 1
OR.vector1.3 <- exp(coef(model3_cov1))
CI.vector1.3 <- exp(confint(model3_cov1))

p.values1.3 <- list(summary(model3_cov1)$coefficients[,4])

#Model 1 cov 2
OR.vector2.3 <- exp(coef(model3_cov2))
CI.vector2.3 <- exp(confint(model3_cov2))

p.values2.3 <- list(summary(model3_cov2)$coefficients[,4])

#Model 1 cov 3
OR.vector3.3.3 <- exp(coef(model3_cov3))
CI.vector3.3.3 <- exp(confint(model3_cov3))

p.values3.3.3 <- list(summary(model3_cov3)$coefficients[,4])

#Model 1 cov all
OR.vector4.3 <- exp(coef(model3_all))
CI.vector4.3 <- exp(confint(model3_all))

p.values4.3 <- list(summary(model3_all)$coefficients[,4])

```


```{r}
#Plotting

stargazer(model3, model3_cov1, model3_cov2,model3_cov3, model3_all, coef = list(OR.vector_m3, OR.vector1.3, OR.vector2.3, OR.vector3.3.3, OR.vector4.3), 
          title = "Table 6. Stepwise model estimates school dropouts and mothers´ marital status, estimates in odd ratios",
          dep.var.labels = "School dropouts",
          ci = T,
ci.custom = list(CI.vector_m3, CI.vector1.3, CI.vector2.3, CI.vector3.3.3, CI.vector4.3), single.row = F, type = "text", out = "model3.htm",
notes= c("Reference category Marital status: Married. Reference category Gender: Male. Reference type of residence: Urban.Reference category Region: Atlantic. Source: DHS Colombia 2015, own estimates"),
 covariate.labels = c("Marital status-Living with partner",
                               "Marital status-Single",
                               "Marital stauts-Other",
                               "Gender-Female",
                               "Type residence-Rural",
                               "Region-Oriental",
                               "Region-Bogota",
                               "Region-Central",
                               "Region-Pacific",
                               "Region-Amazonia"),
          notes.align = "l",
          notes.append = FALSE,
p=
c(p.values_m3, p.values1.3, p.values2.3, p.values3.3.3, p.values4.3))



